Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-02-28 src/tex/ms.tex
[showyourwork] overleaf sync

In the left panel of Fig.~\ref{fig:sbc} we show the distribution of ranks (depicted in red) in each dimension. Highlighted in grey, you see the 99\% confidence interval of a uniform distribution given the number of samples provided. In plain english: for a uniform distribution, we would expect 1 out of 100 (red) bars to lie outside the grey area. This figure shows that overall our posteriors are decently calibrated. Only for the parameter $\log_{10}\left(\mathrm{SFE}\right)$ we see a single bar outside the 99\% confidence interval. But most importantly for the parameters of interest here, $\log_{10}\left(N_\mathrm{Ia}\right)$ and $\alpha_\mathrm{IMF}$ we have a well calibrated posterior.

@@

For each given coverage level $\alpha$, one can then calculate the corresponding average counts and check, whether they correspond to the given $\alpha$. The visualization and interpretation of TARP values is therefore similar to that of SBC. However, in contrast to SBC, TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators. In Fig.~\ref{fig:tarp} we show the result for our NPE in blue in comparison to the ideal line shown in black dashed style. This figure clearly shows that our NPE is well calibrated.

@@

>>>>

In the top panel of Fig.~\ref{fig:tarp} we show the distribution of ranks (depicted in red) in each dimension. Highlighted with black lines, you see the 99\% confidence interval of a uniform distribution given the number of samples provided. In plain english: for a uniform distribution, we would expect 1 out of 100 (blue) bars to lie outside the grey area. This figure shows that overall our posteriors are decently calibrated. Only for the parameter $\log_{10}\left(\mathrm{SFE}\right)$ and Time we see a slight bmiss-calibration. But most importantly for the parameters of interest here, $\log_{10}\left(N_\mathrm{Ia}\right)$ and $\alpha_\mathrm{IMF}$ we have a well calibrated posterior.

@@

For each given coverage level $\alpha$, one can then calculate the corresponding average counts and check, whether they correspond to the given $\alpha$. The visualization and interpretation of TARP values is therefore similar to that of SBC. However, in contrast to SBC, TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators. In the middle row of Fig.~\ref{fig:tarp} we show the result for our NPE in blue in comparison to the ideal line shown in black dashed style. This figure clearly shows that our NPE is well calibrated.

@@

Finally, the bottom row of Fig.~\ref{fig:tarp} shows the predicted vs. true parameter plots where blue dots show the average of the posterior samples and errorbars show the standard deviation of the samples vs their ground truth parameter. We find that for the global parameters we recover the 1:1 relation well while for the other parameters the agreement is tilted.
